---
title: "nidhi"
output: pdf_document
date: "2023-11-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r ordering-competitions-women}

unique(df_women$Competition)

df_women <- df_women %>%
  mutate(competition_number = case_when(
    Competition == "2022 Cottbus World Cup" ~ "1",
    Competition == "2022 Winter Cup" ~ "2",
    Competition == "2022 Doha World Cup" ~ "3",
    Competition == "2022 Cairo World Cup" ~ "4",
    Competition == "2022 British Gymnastics Championships" ~ "5",
    Competition == "2022 Varna World Challenge Cup" ~ "6",
    Competition == "2022 Baku World Cup" ~ "7",
    Competition == "2022 Osijek World Challenge Cup" ~ "8",
    Competition == "2022 9th Senior Artistic Gymnastics Asian Championships" ~ "9",
    Competition == "2022 Koper World Challenge Cup" ~ "10",
    Competition == "2022 U.S. Classic" ~ "11",
    Competition == "BIRMINGHAM 2022 Commonwealth Games" ~ "12",
    Competition == "2022 Senior European Championships" ~ "13",
    Competition == "2022 U.S. Championships" ~ "14",
    Competition == "2022 Paris World Challenge Cup" ~ "15",
    Competition == "2022 Szombathely World Challenge Cup" ~ "16",
    Competition == "2022 Mersin World Challenge Cup" ~ "17",
    Competition == "2022 51st FIG Artistic Gymnastics World Championships" ~ "18",
    Competition == "2023 Cottbus World Cup" ~ "19",
    Competition == "2023 Winter Cup" ~ "20",
    Competition == "2023 Doha World Cup" ~ "21",
    Competition == "2023 Baku World Cup" ~ "22",
    Competition == "EnBW DTB Pokal Team Challenge 2023" ~ "23",
    Competition == "2023 British Gymnastics Championships" ~ "24",
    Competition == "2023 Senior European Championships" ~ "25",
    Competition == "2023 Cairo World Cup" ~ "26",
    Competition == "2023 Varna World Challenge Cup" ~ "27",
    Competition == "2023 Artistic Gymnastics Senior Pan American Championships" ~ "28",
    Competition == "2023 Tel Aviv World Challenge Cup" ~ "29",
    Competition == "2023 Osijek World Challenge Cup" ~ "30",
    Competition == "2023 10th Senior Artistic Gymnastics Asian Championships" ~ "31",
    Competition == "2023 Central American and Caribbean Games" ~ "32",
    Competition == "2023 FISU World University Games" ~ "33",
    Competition == "2023 Core Hydration Classic" ~ "34",
    Competition == "2023 U.S. Championships" ~ "35",
    Competition == "2023 52nd FIG Artistic Gymnastics World Championships" ~ "36",
    TRUE ~ "Other"
  ))

df_women = df_women %>%
  mutate(competition_number = as.numeric(competition_number))

```

```{r categorizing-competitions-women}

df_women <- df_women %>%
  mutate(competition_scope = case_when(
    Competition == "2022 Cottbus World Cup" ~ "World",
    Competition == "2022 Winter Cup" ~ "USA",
    Competition == "2022 Doha World Cup" ~ "World",
    Competition == "2022 Cairo World Cup" ~ "World",
    Competition == "2022 British Gymnastics Championships" ~ "non-USA",
    Competition == "2022 Varna World Challenge Cup" ~ "World",
    Competition == "2022 Baku World Cup" ~ "World",
    Competition == "2022 Osijek World Challenge Cup" ~ "World",
    Competition == "2022 9th Senior Artistic Gymnastics Asian Championships" ~ "non-USA",
    Competition == "2022 Koper World Challenge Cup" ~ "World",
    Competition == "2022 U.S. Classic" ~ "USA",
    Competition == "BIRMINGHAM 2022 Commonwealth Games" ~ "non-USA",
    Competition == "2022 Senior European Championships" ~ "non-USA",
    Competition == "2022 U.S. Championships" ~ "USA",
    Competition == "2022 Paris World Challenge Cup" ~ "World",
    Competition == "2022 Szombathely World Challenge Cup" ~ "World",
    Competition == "2022 Mersin World Challenge Cup" ~ "World",
    Competition == "2022 51st FIG Artistic Gymnastics World Championships" ~ "World",
    Competition == "2023 Cottbus World Cup" ~ "World",
    Competition == "2023 Winter Cup" ~ "USA",
    Competition == "2023 Doha World Cup" ~ "World",
    Competition == "2023 Baku World Cup" ~ "World",
    Competition == "EnBW DTB Pokal Team Challenge 2023" ~ "World",
    Competition == "2023 British Gymnastics Championships" ~ "non-USA",
    Competition == "2023 Senior European Championships" ~ "non-USA",
    Competition == "2023 Cairo World Cup" ~ "World",
    Competition == "2023 Varna World Challenge Cup" ~ "World",
    Competition == "2023 Artistic Gymnastics Senior Pan American Championships" ~ "USA",
    Competition == "2023 Tel Aviv World Challenge Cup" ~ "World",
    Competition == "2023 Osijek World Challenge Cup" ~ "World",
    Competition == "2023 10th Senior Artistic Gymnastics Asian Championships" ~ "non-USA",
    Competition == "2023 Central American and Caribbean Games" ~ "non-USA",
    Competition == "2023 FISU World University Games" ~ "USA",
    Competition == "2023 Core Hydration Classic" ~ "USA",
    Competition == "2023 U.S. Championships" ~ "USA",
    Competition == "2023 52nd FIG Artistic Gymnastics World Championships" ~ "World",
    TRUE ~ "Other"
  ))

```

```{r ordering-competitions-men}

unique(df_men$Competition)

df_men <- df_men %>%
  mutate(competition_number = case_when(
    Competition == "2022 Cottbus World Cup" ~ "1",
    Competition == "2022 Winter Cup" ~ "2",
    Competition == "2022 Doha World Cup" ~ "3",
    Competition == "2022 Cairo World Cup" ~ "4",
    Competition == "2022 British Gymnastics Championships" ~ "5",
    Competition == "2022 Varna World Challenge Cup" ~ "6",
    Competition == "2022 Baku World Cup" ~ "7",
    Competition == "2022 Osijek World Challenge Cup" ~ "8",
    Competition == "2022 9th Senior Artistic Gymnastics Asian Championships" ~ "9",
    Competition == "2022 Koper World Challenge Cup" ~ "10",
    Competition == "2022 U.S. Classic" ~ "11",
    Competition == "BIRMINGHAM 2022 Commonwealth Games" ~ "12",
    Competition == "2022 Senior European Championships" ~ "13",
    Competition == "2022 U.S. Championships" ~ "14",
    Competition == "2022 Paris World Challenge Cup" ~ "15",
    Competition == "2022 Szombathely World Challenge Cup" ~ "16",
    Competition == "2022 Mersin World Challenge Cup" ~ "17",
    Competition == "2022 51st FIG Artistic Gymnastics World Championships" ~ "18",
    Competition == "2023 Cottbus World Cup" ~ "19",
    Competition == "2023 Winter Cup" ~ "20",
    Competition == "2023 Doha World Cup" ~ "21",
    Competition == "2023 Baku World Cup" ~ "22",
    Competition == "EnBW DTB Pokal Team Challenge 2023" ~ "23",
    Competition == "2023 British Gymnastics Championships" ~ "24",
    Competition == "2023 Senior European Championships" ~ "25",
    Competition == "2023 Cairo World Cup" ~ "26",
    Competition == "Oceania Continental Championships 2023" ~ "27",
    Competition == "2023 Varna World Challenge Cup" ~ "28",
    Competition == "2023 Artistic Gymnastics Senior Pan American Championships" ~ "29",
    Competition == "2023 Tel Aviv World Challenge Cup" ~ "30",
    Competition == "2023 Osijek World Challenge Cup" ~ "31",
    Competition == "2023 10th Senior Artistic Gymnastics Asian Championships" ~ "32",
    Competition == "2023 Central American and Caribbean Games" ~ "33",
    Competition == "2023 FISU World University Games" ~ "34",
    Competition == "2023 Core Hydration Classic" ~ "35",
    Competition == "2023 U.S. Championships" ~ "36",
    Competition == "2023 52nd FIG Artistic Gymnastics World Championships" ~ "37",
    TRUE ~ "Other"
  ))

df_men = df_men %>%
  mutate(competition_number = as.numeric(competition_number))

```

```{r categorizing-competitions-men}

df_men <- df_men %>%
  mutate(competition_scope = case_when(
    Competition == "2022 Cottbus World Cup" ~ "World",
    Competition == "2022 Winter Cup" ~ "USA",
    Competition == "2022 Doha World Cup" ~ "World",
    Competition == "2022 Cairo World Cup" ~ "World",
    Competition == "2022 British Gymnastics Championships" ~ "non-USA",
    Competition == "2022 Varna World Challenge Cup" ~ "World",
    Competition == "2022 Baku World Cup" ~ "World",
    Competition == "2022 Osijek World Challenge Cup" ~ "World",
    Competition == "2022 9th Senior Artistic Gymnastics Asian Championships" ~ "non-USA",
    Competition == "2022 Koper World Challenge Cup" ~ "World",
    Competition == "2022 U.S. Classic" ~ "USA",
    Competition == "BIRMINGHAM 2022 Commonwealth Games" ~ "non-USA",
    Competition == "2022 Senior European Championships" ~ "non-USA",
    Competition == "2022 U.S. Championships" ~ "USA",
    Competition == "2022 Paris World Challenge Cup" ~ "World",
    Competition == "2022 Szombathely World Challenge Cup" ~ "World",
    Competition == "2022 Mersin World Challenge Cup" ~ "World",
    Competition == "2022 51st FIG Artistic Gymnastics World Championships" ~ "World",
    Competition == "2023 Cottbus World Cup" ~ "World",
    Competition == "2023 Winter Cup" ~ "USA",
    Competition == "2023 Doha World Cup" ~ "World",
    Competition == "2023 Baku World Cup" ~ "World",
    Competition == "EnBW DTB Pokal Team Challenge 2023" ~ "World",
    Competition == "2023 British Gymnastics Championships" ~ "non-USA",
    Competition == "2023 Senior European Championships" ~ "non-USA",
    Competition == "2023 Cairo World Cup" ~ "World",
    Competition == "Oceania Continental Championships 2023" ~ "non-USA",
    Competition == "2023 Varna World Challenge Cup" ~ "World",
    Competition == "2023 Artistic Gymnastics Senior Pan American Championships" ~ "USA",
    Competition == "2023 Tel Aviv World Challenge Cup" ~ "World",
    Competition == "2023 Osijek World Challenge Cup" ~ "World",
    Competition == "2023 10th Senior Artistic Gymnastics Asian Championships" ~ "non-USA",
    Competition == "2023 Central American and Caribbean Games" ~ "non-USA",
    Competition == "2023 FISU World University Games" ~ "USA",
    Competition == "2023 Core Hydration Classic" ~ "USA",
    Competition == "2023 U.S. Championships" ~ "USA",
    Competition == "2023 52nd FIG Artistic Gymnastics World Championships" ~ "World",
    TRUE ~ "Other"
  ))

```


Making Visualizations

Why we aren't considering penalties:

For both men's and women's competitions, the 95th percentile for the penalties is less than 0.5 (it is 0.5 for men and 0.4 for women). This means that 95% of penalties for both genders are not more than 0.5. The visualizations show that this does not change drastically based on the scope of the competition.
Additionally, since the 51st percentile of penalties for both genders is 0.1, this means that the majority of entries had a penalty score of 0.1 or less.
```{r}

# for women
quantile(as.numeric(df_women_c$Penalty), 0.95, na.rm = TRUE)
quantile(as.numeric(df_women_c$Penalty), 0.51, na.rm = TRUE)

ggplot(data = df_women_c, aes(x = Penalty)) +
  geom_histogram(bins = 30) +
  facet_wrap(~ competition_scope) +
  labs(title = "Penalty Distribution for Women Athletes",
       x = "Penalty",
       y = "Count")

# for men
quantile(as.numeric(df_men_c$Penalty), 0.95, na.rm = TRUE)
quantile(as.numeric(df_men_c$Penalty), 0.51, na.rm = TRUE)

ggplot(data = df_men_c, aes(x = Penalty)) +
  geom_histogram(bins = 30) +
  facet_wrap(~ competition_scope) +
    labs(title = "Penalty Distribution for Men Athletes",
       x = "Penalty",
       y = "Count")
```

Why we're using mean for bootstrapping:

The distribution of event scores for women and men are both approximately normally distributed, with the mean and median scores less than 0.5 apart within each gender. 
The mean of women's scores is 12.207, and the median of women's scores is 12.367.
The mean of men's scores is 13.048, and the median of men's scores is 13.233.
```{r}
ggplot(data = df_women_c, aes(x = Score)) +
  geom_histogram() +
  labs(title = "Score Distribution for Women Athletes",
       x = "Score",
       y = "Count")

ggplot(data = df_men_c, aes(x = Score)) +
  geom_histogram() +
  labs(title = "Score Distribution for Men Athletes",
       x = "Score",
       y = "Count")

mean(df_women_c$Score)
median(df_women_c$Score)

mean(df_men_c$Score)
median(df_men_c$Score)
```

Whether scores vary by type of round in competition:
Doesn't seem to vary much; make the assumption that qualification scores are representative of an athlete's scores in the medal rounds

```{r}
ggplot(data = df_women_c, aes(x = Score)) +
  geom_histogram(aes(fill = Round)) +
  labs(title = "Score Distribution for Women Athletes, by Round",
       x = "Score",
       y = "Count")

ggplot(data = df_men_c, aes(x = Score)) +
  geom_histogram(aes(fill = Round)) +
  labs(title = "Score Distribution for Men Athletes, by Round",
       x = "Score",
       y = "Count")

mean(df_women_c$Score)
median(df_women_c$Score)

mean(df_men_c$Score)
median(df_men_c$Score)
```
Execution scores have a higher range than difficulty scores, for both men and women athletes (colored by round but doesn't seem to be much difference based on round)

```{r}

ggplot(data = df_men_c, aes(x = D_Score)) +
  geom_histogram(aes(fill = Round)) +
  labs(title = "Difficulty Score Distribution for Men Athletes",
       x = "Score",
       y = "Count")

ggplot(data = df_men_c, aes(x = E_Score)) +
  geom_histogram(aes(fill = Round)) +
  labs(title = "Execution Score Distribution for Men Athletes",
       x = "Score",
       y = "Count")


```
```{r}
ggplot(data = df_women_c, aes(x = D_Score)) +
  geom_histogram(aes(fill = Round)) +
  labs(title = "Difficulty Score Distribution for Women Athletes",
       x = "Score",
       y = "Count")

ggplot(data = df_women_c, aes(x = E_Score)) +
  geom_histogram(aes(fill = Round)) +
  labs(title = "Execution Score Distribution for Women Athletes",
       x = "Score",
       y = "Count")
```
```{r}

ggplot(data = df_women_c, aes(x = Score)) +
  geom_histogram(aes(fill = Country)) +
  labs(title = "Score Distribution by Country",
       x = "Score",
       y = "Count")

ggplot(data = df_men_c, aes(x = Score)) +
  geom_histogram(aes(fill = Country)) +
  labs(title = "Score Distribution by Country",
       x = "Score",
       y = "Count")

```
Do the above visualization later for only the "important" versus "non-important" countries once we find them out from the simulation

